---
title: "AutocorrExperimentStats"
output:
  html_document: default
  pdf_document: default
date: "2025-11-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is the code for fitting a bias-reduced linear regression model to the experimental extinction data, then calculating confidence intervals, McFadden's pseudo R\^2, and comparing residual with null deviance:

```{r modelfit}
data <- read.csv("Rdata.csv")

install.packages("brglm")
library(brglm)

#Model fit
model <- brglm(formula = Extinction ~ Mu * Gamma, family = "binomial", data = data)
summary(model)

#Confidence intervals
confint.default(model)

#McFadden's pseudo R^2
(1-model$deviance/model$null.deviance)*100

#deviance comparison
deviance(model)/df.residual(model)
```

## Including Plots

This code creates Fig 5, the plot of the brglm model fit compared to experimental results:

```{r plot, echo=TRUE}
library(ggplot2)
library(dplyr)

data$Gamma <- factor(data$Gamma)
data$Mu <- as.numeric(data$Mu)

model <- glm(Extinction ~ Mu * Gamma, data = data, family = "binomial")

predicted_probs <- predict(model, type = "response")
plot_data <- data.frame(
  Mu = data$Mu,
  Gamma = factor(data$Gamma),
  observed = data$Extinction,
  predicted = predicted_probs
)

summary_data <- data %>%
  group_by(Mu, Gamma) %>%
  summarise(
    prop_extinct = mean(Extinction),
    n = n(),
    se = sqrt(prop_extinct * (1 - prop_extinct) / n),
    ci_lower = prop_extinct - 1.96 * se,
    ci_upper = prop_extinct + 1.96 * se
  ) %>%
  ungroup()

summary_data$Gamma <- factor(summary_data$Gamma)
plot_data$Gamma <- factor(plot_data$Gamma)

ggplot() +
  geom_line(data = plot_data,
            aes(x = Mu, y = predicted, color = Gamma, group = Gamma),
            size = 1.2) +
  geom_point(data = summary_data,
             aes(x = Mu, y = prop_extinct, color = Gamma),
             size = 3) +
  geom_errorbar(data = summary_data,
                aes(x = Mu, ymin = ci_lower, ymax = ci_upper, color = Gamma),
                width = 0.2) +
  labs(title = "Observed vs Predicted Persistence Probabilities",
       x = "Mean Temperature",
       y = "Probability of Persistence",
       color = "Gamma") +
  theme_minimal() +
  scale_color_manual(values = c("gray50", "pink2", "firebrick3"))
```

Sample of how we ran X^2 tests (showing Table S7a: strong boundary, all treatments). To run a different test, switch in the data from the desired contingency table (tables for all computed tests found in supplemental material S5).

```{r chi squared}
data <- matrix(c(60,3,48,69),nrow=2)
result <- chisq.test(data)
print(result)
```

Code for conducting post-hoc pairwise comparison. The model fit is slightly adjusted here: in order to conduct this test, we needed to compare within mean temperature treatments. We thus used 'ExactMu' (the exact goal mean temperature for each treatment) for the mean temperature instead of 'Mu' (the measured mean temperature of each treatment) and treated mean temperature as a factor.

```{r post-hoc pairwise}
data <- read.csv("Rdata.csv")
data$ExactMu <- as.factor(data$ExactMu)
data$Gamma <- as.factor(data$Gamma)

model <- brglm(formula = Extinction ~ ExactMu * Gamma, family = "binomial", data = data)
emm <- emmeans(model, ~ Gamma | ExactMu, type = "response")
pairs(emm, adjust = "tukey")
```