---
title: "rTPCs"
output:
  html_document: default
  pdf_document: default
date: "2024-06-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitting TPCs from rTPC

This document's purpose is to fit TPCs for experimental cultures that survived various temperature regimes.

```{r, libraries, message = FALSE}
# R libraries needed
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)


```

```{r, message = FALSE}
#read in data
tpc_data_all<-read_csv("post_experiment_tpc_alison_0525.csv")
tpc_data_all<-rename(tpc_data_all,Temperature=`Incubation Temperature (°C)`)
tpc_data_all<-rename(tpc_data_all,Daily_growth_rate=`PGR (r) (per day)`)
tpc_data_all$Daily_growth_rate=as.numeric(tpc_data_all$Daily_growth_rate)
tpc_data_all$Source<-as.factor(tpc_data_all$Source)
```


```{r}
#graph data
ggplot(tpc_data_all) +
  geom_point( aes(x=Temperature, y=Daily_growth_rate,color=Source))+
  labs(x = 'Temperature (ºC)',
       y = 'Population growth rate (r, per day)',
       title = 'P. caudatum growth rate across temperatures')

```
#Lactin2 model

For this analysis we are using the lactin2 model on all 6 protist culture sources


```{r}
#lactin2_1995

  # choose model
  low_extend<-7
  high_extend<-1
  mod = "lactin2_1995"
  message("\nFUNCTION STRUCTURE lactin2_1995:")
lactin2_1995
```

for the source culture
```{r}
### culture by culture ##### 

# for the source culture
tpc_data<-tpc_data_all %>% filter(Source=="stock") %>% na.omit()
# get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit_stock <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit_stock) %>%
    # round for easy viewing
    mutate_all(round, 2)
  message("MODEL FIT stock:")
fit_stock
summary(fit_stock)
glance(fit_stock)

CI <- nlstools::confint2(fit_stock) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI

###############
# source = W1-10
# for the source culture
tpc_data<-tpc_data_all %>% filter(Source=="W1-10") %>% na.omit()
# get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit_W1_10 <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit_W1_10) %>%
    # round for easy viewing
    mutate_all(round, 2)
  message("MODEL FIT W1-10:")
fit_W1_10
summary(fit_W1_10)
glance(fit_W1_10)

CI <- nlstools::confint2(fit_W1_10) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI
###############
# source = W1-10
# for the source culture
tpc_data<-tpc_data_all %>% filter(Source=="W1-10") %>% na.omit()
# get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit_W1_10 <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit_W1_10) %>%
    # round for easy viewing
    mutate_all(round, 2)
  message("MODEL FIT W1-10:")
fit_W1_10
summary(fit_W1_10)
glance(fit_W1_10)

CI <- nlstools::confint2(fit_W1_10) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI

###############
# source = W4-6
# for the source culture
tpc_data<-tpc_data_all %>% filter(Source=="W4-6") %>% na.omit()
# get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit_W4_6 <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit_W4_6) %>%
    # round for easy viewing
    mutate_all(round, 2)
  message("MODEL FIT W1-10:")
fit_W4_6
summary(fit_W4_6)
glance(fit_W4_6)

CI <- nlstools::confint2(fit_W4_6) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI


###############
# source = P1-9
# for the source culture
tpc_data<-tpc_data_all %>% filter(Source=="P1-9") %>% na.omit()
# get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit_P1_9 <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit_P1_9) %>%
    # round for easy viewing
    mutate_all(round, 2)
  message("MODEL FIT W1-10:")
fit_P1_9
summary(fit_P1_9)
glance(fit_P1_9)

CI <- nlstools::confint2(fit_P1_9) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI

###############
# source = P4-2
# for the source culture
tpc_data<-tpc_data_all %>% filter(Source=="P4-2") %>% na.omit()
# get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit_P4_2 <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit_P4_2) %>%
    # round for easy viewing
    mutate_all(round, 2)
  message("MODEL FIT W1-10:")
fit_P4_2
summary(fit_P4_2)
glance(fit_P4_2)

CI <- nlstools::confint2(fit_P4_2) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI


###############
# source = B1-6
# for the source culture
tpc_data<-tpc_data_all %>% filter(Source=="B1-6") %>% na.omit()
# get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit_B1_6 <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit_B1_6) %>%
    # round for easy viewing
    mutate_all(round, 2)
  message("MODEL FIT B1-6:")
fit_B1_6
summary(fit_B1_6)
glance(fit_B1_6)

CI <- nlstools::confint2(fit_B1_6) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI

```

```{r}
# predict new data
  new_data <- data.frame(Temperature = seq(min(tpc_data$Temperature)-low_extend, max(tpc_data$Temperature)+high_extend, 0.25))
  preds <- augment(fit_stock, newdata = new_data %>% select(Temperature))
  new_data$stock<-preds$.fitted
  preds <- augment(fit_W1_10, newdata = new_data %>% select(Temperature))
  new_data$`W1-10`<-preds$.fitted
  preds <- augment(fit_W4_6, newdata = new_data %>% select(Temperature))
  new_data$`W4-6`<-preds$.fitted
  preds <- augment(fit_P1_9, newdata = new_data %>% select(Temperature))
  new_data$`P1-9`<-preds$.fitted
  preds <- augment(fit_P4_2, newdata = new_data %>% select(Temperature))
  new_data$`P4-2`<-preds$.fitted
  preds <- augment(fit_B1_6, newdata = new_data %>% select(Temperature))
  new_data$`B1-6`<-preds$.fitted
  new_data<-new_data %>% pivot_longer(c("stock","W1-10","W4-6","P1-9","P4-2","B1-6"), names_to="Source",values_to="Daily_growth_rate")
  new_data$Source<-as.factor(new_data$Source)
  # plot data and model fit
  ggplot() +
    geom_point(aes(Temperature, Daily_growth_rate,color = Source),tpc_data_all) +
    geom_line(aes(Temperature, Daily_growth_rate,color = Source), new_data) +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'lactin2_1995 fit')
  
  ggplot() +
    geom_point(aes(Temperature, Daily_growth_rate,color = Source),tpc_data_all%>% filter(Source=="stock")) +
    geom_line(aes(Temperature, Daily_growth_rate,color = Source), new_data %>% filter(Source=="stock")) +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'stock lactin2_1995 fit')
  
    ggplot() +
    geom_point(aes(Temperature, Daily_growth_rate,color = Source),tpc_data_all%>% filter(Source=="W1-10")) +
    geom_line(aes(Temperature, Daily_growth_rate,color = Source), new_data %>% filter(Source=="W1-10")) +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'W1-10 lactin2_1995 fit')
    
      ggplot() +
    geom_point(aes(Temperature, Daily_growth_rate,color = Source),tpc_data_all%>% filter(Source=="W4-6")) +
    geom_line(aes(Temperature, Daily_growth_rate,color = Source), new_data %>% filter(Source=="W4-6")) +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'W4-6 lactin2_1995 fit')
      
        ggplot() +
    geom_point(aes(Temperature, Daily_growth_rate,color = Source),tpc_data_all%>% filter(Source=="P1-9")) +
    geom_line(aes(Temperature, Daily_growth_rate,color = Source), new_data %>% filter(Source=="P1-9")) +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'P1_9 lactin2_1995 fit')
        
          ggplot() +
    geom_point(aes(Temperature, Daily_growth_rate,color = Source),tpc_data_all%>% filter(Source=="P4-2")) +
    geom_line(aes(Temperature, Daily_growth_rate,color = Source), new_data %>% filter(Source=="P4-2")) +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'P4-2 lactin2_1995 fit')
          
    ggplot() +
    geom_point(aes(Temperature, Daily_growth_rate,color = Source),tpc_data_all%>% filter(Source=="B1-6")) +
    geom_line(aes(Temperature, Daily_growth_rate,color = Source), new_data %>% filter(Source=="B1-6")) +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'B1-6 lactin2_1995 fit')
```








# OLD DATA for comparison 

```{r, message = FALSE}
#read in data
tpc_data<-read_csv("TPC DATA caudatum 0617.csv")
tpc_data<-rename(tpc_data,Temperature=`Incubation Temperature (°C)`)
tpc_data<-rename(tpc_data,Daily_growth_rate=`PGR (r) (per day)`)
```




### lactin2_1995

This model produces negative growth rates on both ends and looks reasonable overall--worth a look. It's not piecewise. 

```{r}
#lactin2_1995
  
  # choose model
  low_extend<-7
  high_extend<-3
  mod = "lactin2_1995"
  # get start vals
  start_vals <- get_start_vals(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # get limits
  low_lims <- get_lower_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  upper_lims <- get_upper_lims(tpc_data$Temperature, tpc_data$Daily_growth_rate, model_name = mod)
  
  # fit model
  fit <- nls_multstart(Daily_growth_rate~lactin2_1995(temp = Temperature, 
                                                  a, b, tmax, delta_t),
                       data = tpc_data,
                       iter = 500,
                       start_lower = start_vals - 10,
                       start_upper = start_vals + 10,
                       lower = low_lims,
                       upper = upper_lims,
                       supp_errors = 'Y')
  
  # calculate additional traits
  other_traits<-calc_params(fit) %>%
    # round for easy viewing
    mutate_all(round, 2)

  
  # predict new data
  new_data <- data.frame(Temperature = seq(min(tpc_data$Temperature)-low_extend, max(tpc_data$Temperature)+high_extend, 0.25))
  preds <- augment(fit, newdata = new_data)
  
  # plot data and model fit
  ggplot(tpc_data, aes(Temperature, Daily_growth_rate)) +
    geom_point() +
    geom_line(aes(Temperature, .fitted), preds, col = 'blue') +
    theme_bw(base_size = 12) +
    labs(x = 'Temperature (ºC)',
         y = 'PGR',
         title = 'lactin2_1995 fit')
```

```{r echo=FALSE}
#lactin2_1995 fit
message("MODEL FIT lactin2_1995:")
fit
summary(fit)
library(broom)
glance(fit)
tidy(fit)


CI <- nlstools::confint2(fit) %>%
  data.frame() %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..)
CI

message("\nFUNCTION STRUCTURE lactin2_1995:")
lactin2_1995
```



 
